CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-10-01 codecheck-paper.tex
use only DOI of references, make DOIs clickable in PDF
update description of register and certificate

\usepackage{booktabs}

@@

\section*{Register}\label{register}

We have a curated list of 15 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons (do we name them all?)
from interesting historical papers, to papers codechecked as part of
peer review, or because of public interest (covid). several covid models
have been done during the time of peer review of these papers, under our
initative rather than requested from journal, but certificate have then
been acknowledged/cited in the paper (LANCET / LSHTM).
Table~\ref{tab:register} lists the certificates completed to date,
along with citations to the original papers.

The bulk of the certificates fall into three themes: (1) `classic'
papers from computational neuroscience (2) Covid modelling preprints, and
(3) GIScience.  
The first theme was the initial set of papers used to explore the concept
of CODECHECK. The idea was to take very well known articles from a specific
domain, in this case the work area of co-author S.~Eglen, possibly as as a
conversation starter for potential collaborating journals.
The second theme is rather opportunistic. As the COVID-19 pandemic swept
over the whole Earth, we hoped to contribute to science's answer to the
unknown virus. The checks were solicited through community interaction.
The third theme represents co-author D.~N\"ust's work area.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series. % TODO add reference to MUNIN talk
While using slightly different terms and infrastructue, e.g., `reproducibility
reports' are published on OSF, instead of certificates on Zenodo, the 
reproducibility reviews adhere to the codecheck principles.
Furthermore, two checks were completed as part of a peer review for a
GIScience journal.

Out of these areas, we would like to briefly highlight three examples here.

@@

  \begin{tabular}{llp{11cm}}
    \toprule
    Certificate & Research area & Description \\ \midrule
    2020-001 \cite{cert-2020-001} &
    Machine learning & CODECHECK undertaken post acceptance of
    manuscript and before its publication in \textit{Gigascience}
    \cite{Piccolo2020-lo}.
    \\
    2020-002  \cite{cert-2020-002} & neuroscience & Paper from 1997
 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
 Code written for this project.
    \\
    2020-003  \cite{cert-2020-003} & neuroscience &
                                                     Classic paper
                                                     from 1982 on
                                                     models of
                                                     associative
                                                     memory
                                                     \cite{Hopfield1982-mz}.
    (Simulation code was written specifically for this project.)\\
    2020-004  \cite{cert-2020-004} & area & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
    2020-005  \cite{cert-2020-005} & area & Larisch \\
    2020-006  \cite{cert-2020-006} & area & Detorakis \\
    2020-007  \cite{cert-2020-007} & area & in progress \\
    2020-008  \cite{cert-2020-008} & Covid & Modelling of
    interventions on Covid  cases in the UK; checked as preprint (NOW REMOVED XXX)
    and now published \cite{Davies2020-vj} \\
    2020-009  \cite{cert-2020-009} & Covid &  tracing \\
    2020-010  \cite{cert-2020-010} & Covid &  report 9 \\
    2020-011  \cite{cert-2020-011} & Covid & Modelling of Covid
    spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
    2020-012  \cite{cert-2020-012} & Covid &   Modelling of Covid spread across the USA; preprint currently review \cite{Unwin2020}. \\
    2020-013  \cite{cert-2020-013} & area & biorXiv \cite{Spitschan2020.06.02.129502}. \\
    2020-014  \cite{cert-2020-014} & area & eLife \cite{Sadeh2020}. \\
    2020-015  \cite{cert-2020-015} & area & eLife \cite{Liou2020}. \\
    2020-016  \cite{cert-2020-016} & GIScience & 
    R models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020}. \\
    2020-017  \cite{cert-2020-017} & GIScience & 
    Spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020}. \\
    2020-018  \cite{cert-2020-018} & GIScience & 
    Cellular automaton in R for modeling dynamic phenomena within a Discrete Global Grid System (DGGS) data model; using a subset for demonstration \cite{Hojati2020}. \\
    2020-019  \cite{cert-2020-019} & GIScience & Reachability analysis of suburban transportation using analysis of privately owned shared cars; workflow uses Python, and reproduction with a considerably subsampled dataset \cite{Illium2020}. \\
    2020-020  \cite{cert-2020-020} & GIScience & In-database windows operators for processing spatio-temporal data streams based on Python and PostgreSQL/PostGIS \cite{Werner2020}; reproduction using container. \\
    2020-021  \cite{cert-2020-021} & GIScience & Python code for comparing supervised machine learning models for spatial nominal entity recognition. \cite{Medad2020}. \\
    2020-022  \cite{cert-2020-022} & GIScience & Python code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\
    2020-023  \cite{cert-2020-023} & GIScience & Analysis of spatial footprints of geo-tagged extreme weather events from social media using R \cite{Owuor2020}. \\
    \\ \bottomrule
  \end{tabular}
  \caption{Register of completed certificates as of October 2020.  An interactive version
  of this is available at \url{http://codecheck.org.uk/register}.  SJE
to complete...}
  \label{tab:register}
\end{table}

@@

At the completion of each CODECHECK, a certificate is written that
states which figures and tables from the original article could be
reproduced by the CODECHECKER.  This certificate is made freely
available so that all readers of the article can see which elements
have been reproduced.  The certificate also includes links to the code
and data used by the codechecker, allowing others to build on the
work.  The format of the certificates has evolved slightly during the
project, as we have learnt to automate different aspects of the
certification.

To demonstrate the typical contents of a certificate, in
Figure~\ref{fig:cert} we have annotated the first four (out of ten)
pages of the certificate 2020-012 \cite{cert-2020-012}.  The reference
article \cite{Unwin2020} describes the modelling COVID-19 across the
USA.
%% TODO - has this been published yet?.  No, not yet... in review...
Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
which points to the certificate as archived on Zenodo.  Our logo is
attached to the certificate to denote successful reproduction.
Figure~\ref{fig:cert}B provides the key metadata about the check --
which paper was checked, and who/when it was checked.  Materials
relating to the codecheck are freely available via the Repository
link.  Figure~\ref{fig:cert}C is a summary of how the codecheck was
performed, and any interesting findings.

>>>>

\usepackage{booktabs}
\usepackage{doi} % turn DOIs into links

@@

\section*{Register}\label{register}
% count certificates: 
% curl https://raw.githubusercontent.com/codecheckers/register/master/register.csv | wc -l
% minus 1 for Hathway-Goodman (no empty last line, so no -1 for the headline needed)

We have a curated list of 22 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons and fall into three themes:
(1) `classic' papers from computational neuroscience
(2) COVID-19 modelling preprints, and
(3) GIScience.  
%from interesting historical papers, to papers codechecked as part of
%peer review, or because of public interest (COVID-19).

The first theme was the initial set of papers used to explore the concept
of CODECHECK. The idea was to take very well known historical articles from a 
specific domain, in this case the work area of co-author S.~Eglen, possibly 
as as a conversation starter for potential collaborating journals.
The second theme is rather opportunistic. As the COVID-19 pandemic swept
over the whole Earth, we hoped to contribute a small part to science's answer
to the unknown virus. The checks were solicited through community interaction
%Several COVID-19 models have been done during the time of peer review of these papers,
and under our initative rather than requested from journal, but the certificate
has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
The third theme represents co-author D.~N\"ust's work area.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series. % TODO add reference to MUNIN talk
While using slightly different terms and infrastructue, e.g., `reproducibility
reports' are published on OSF, instead of certificates on Zenodo, the 
reproducibility reviews adhere to the codecheck principles.
Furthermore, two checks were completed as part of a peer review for a
GIScience journal.

Table~\ref{tab:register} lists the certificates completed to date,
along with citations to the original papers.
Out of these areas, we would like to briefly highlight three examples here.

@@

  \begin{tabular}{llp{11cm}}
    \toprule
    Certificate & Research area & Description \\ \midrule
    2020-001 \cite{cert-2020-001} &
    Machine learning & CODECHECK undertaken post acceptance of
    manuscript and before its publication in \textit{Gigascience}
    \cite{Piccolo2020-lo}. \\
    2020-002  \cite{cert-2020-002} & neuroscience & Paper from 1997
 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
 Code written for this project. \\
    2020-003  \cite{cert-2020-003} & neuroscience &  Classic paper
                                                     from 1982 on
                                                     models of
                                                     associative
                                                     memory
                                                     \cite{Hopfield1982-mz}.
    (Simulation code was written specifically for this project.)\\
    2020-004  \cite{cert-2020-004} & area & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
    2020-005  \cite{cert-2020-005} & area & Larisch \\
    2020-006  \cite{cert-2020-006} & area & Detorakis \\
    2020-007  \cite{cert-2020-007} & area & in progress \\
    2020-008  \cite{cert-2020-008} & COVID-19 & Modelling of
    interventions on COVID-19 cases in the UK; checked as preprint (NOW REMOVED XXX)
    and now published \cite{Davies2020-vj} \\
    2020-009  \cite{cert-2020-009} & COVID-19 & tracing \\
    2020-010  \cite{cert-2020-010} & COVID-19 & report 9 \\
    2020-011  \cite{cert-2020-011} & COVID-19 & Modelling of Covid
    spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
    2020-012  \cite{cert-2020-012} & COVID-19 & Modelling of COVID-19 spread across the USA; preprint currently review \cite{Unwin2020}. \\
    2020-013  \cite{cert-2020-013} & area & biorXiv \cite{Spitschan2020.06.02.129502}. \\
    2020-014  \cite{cert-2020-014} & area & eLife \cite{Sadeh2020}. \\
    2020-015  \cite{cert-2020-015} & area & eLife \cite{Liou2020}. \\
    2020-016  \cite{cert-2020-016} & GIScience & 
    R models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020}. \\
    2020-017  \cite{cert-2020-017} & GIScience & 
    Spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020}. \\
    2020-018  \cite{cert-2020-018} & GIScience & 
    Cellular automaton in R for modeling dynamic phenomena within a Discrete Global Grid System (DGGS) data model; using a subset for demonstration \cite{Hojati2020}. \\
    2020-019  \cite{cert-2020-019} & GIScience & Reachability analysis of suburban transportation using analysis of privately owned shared cars; workflow uses Python, and reproduction with a considerably subsampled dataset \cite{Illium2020}. \\
    2020-020  \cite{cert-2020-020} & GIScience & In-database windows operators for processing spatio-temporal data streams based on Python and PostgreSQL/PostGIS \cite{Werner2020}; reproduction using container. \\
    2020-021  \cite{cert-2020-021} & GIScience & Python code for comparing supervised machine learning models for spatial nominal entity recognition. \cite{Medad2020}. \\
    2020-022  \cite{cert-2020-022} & GIScience & Python code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\
    2020-023  \cite{cert-2020-023} & GIScience & Analysis of spatial footprints of geo-tagged extreme weather events from social media using R \cite{Owuor2020}. \\
    \\ \bottomrule
  \end{tabular}
  \caption{Register of completed certificates as of October 2020.  An interactive version
  of this is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
  }
  \label{tab:register}
\end{table}

@@

At the completion of each CODECHECK, a certificate is written that
states which outputs, i.e., numbers, figures and tables, from the 
original article could be reproduced by the CODECHECKER.
This certificate is made openly
available so that all readers of the article can see which elements
have been reproduced.  The certificate also includes links to the code
and data used by the codechecker, allowing others to build on the
work.
The format of the certificates has evolved slightly during the
project, as we have learnt to automate different aspects of the
certification.

To demonstrate the typical contents of a certificate, in
Figure~\ref{fig:cert} we have annotated the first four out of ten
pages of the certificate \texttt{2020-012} \cite{cert-2020-012}.
The referenced article \cite{Unwin2020} describes the modelling COVID-19
across the USA.
%% TODO - has this been published yet?.  No, not yet... in review...
Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
which points to the certificate as archived on Zenodo.
The CODECHECK logo is
attached to the certificate to denote successful reproduction.
Figure~\ref{fig:cert}B provides the key metadata about the check --
which paper was checked, who the authors and the codechecker were,
and when it was checked. Materials
relating to the codecheck are freely available via the Repository
link. This table is generated from machine readable text file, the
CODECHECK configuration file \texttt{codecheck.yml}.
The configuration file collects the core pieces of metadata surrounding
a check and enables current and future automation of workflows and 
meta-analyses.
Figure~\ref{fig:cert}C is a textual summary of how the codecheck was
performed, and any interesting findings.
